---
title: "Different Filters"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
#avalik1 contains data about years 2016-2017
avalik1 <- read.csv2('avalik_1.csv', sep ='\t', encoding = 'UTF-8')
#seventeenonly contains data from
seventeenonly <- filter(avalik1, substr(avalik1$ToimKpv,1,4) == '2017')
#avalik2 contains data about years 2012-2016
avalik2 <- read.csv2('avalik_2.csv', sep ='\t', encoding = 'UTF-8')
```
## Notation
n is often used to denote a number of counted entities.


# Grouping criminal activity by hours

## Year 2017
The misdemeanor reportings having been grouped by the hour, at which they happened gave the following results. 

```{r, echo=FALSE}
#Remove rows with no information about the time
so_new <- data.frame(filter(seventeenonly, !(seventeenonly$ToimKell=='')))
#Grouping misdemeanor by hours
seventeenonly_by_hour <- count(so_new, "Hour" = substr(so_new$ToimKell, 1, 2))
print("Most active hours")
head(seventeenonly_by_hour[order(-seventeenonly_by_hour$n),], 4)
print("Least active hours")
head(seventeenonly_by_hour[order(seventeenonly_by_hour$n),], 4)
```
A graphical visualisation displays the following:

```{r, echo=FALSE}
#Grouping misdemeanor by hours
seventeenonly_hour_rate_t <- prop.table(table("Hour" = substr(so_new$ToimKell, 1, 2))) 
seventeenonly_hour_rate <- data.frame(seventeenonly_hour_rate_t, row.names = names(seventeenonly_hour_rate_t))
```

```{r, echo=FALSE}
ggplot(data=seventeenonly_hour_rate, aes(x=Hour, y=Freq)) +
   geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
   guides(fill=FALSE) +
   xlab("Hour") + ylab("Probability of a misdemeanor happening") +
   ggtitle("Distribution of reported cases during the day")
```

## Previous years
The misdemeanor reportings having been grouped by the hour, at which they happened gave the following results. 

```{r, echo=FALSE}
#Remove rows with no information about the time
avalik2_new <- data.frame(filter(avalik2, !(avalik2$ToimKell=='')))
#Group by hour
avalik2_by_hour <- count(avalik2_new, "Hour" = substr(ToimKell, 1, 2))
print("Most active hours")
head(avalik2_by_hour[order(-avalik2_by_hour$n),], 4)
print("Least active hours")
head(avalik2_by_hour[order(avalik2_by_hour$n),],4)
```

A graphical visualisation displays the following:

```{r, echo=FALSE}
avalik2_hour_rate_t <- prop.table(table("Hour" = substr(avalik2_new$ToimKell, 1, 2)))
avalik2_hour_rate <- data.frame(avalik2_hour_rate_t, row.names = names(avalik2_hour_rate_t))
```

```{r, echo=FALSE}
ggplot(data=avalik2_hour_rate, aes(x=Hour, y=Freq)) +
   geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
   guides(fill=FALSE) +
   xlab("Hour") + ylab("Probability of a misdemeanor happening") +
   ggtitle("Distribution of reported cases during the day 2012-2016")

```

## Conclusion

The most active hours are around 17 o'clock. The least are at 6 in the morning.
There have been no major changes to the distribution of the rate of criminal activity.

# Grouping by type of misdemeanor
In the original data, each row contained one or more classifiers. 
## Year 2017

```{r, echo=FALSE}
misdemeanor_types_all <- unlist(strsplit(as.character(seventeenonly$SyndmusLiik), ","))
misdemeanor_types <- unique(misdemeanor_types_all)

table_of_misdemeanortype_freq <- table("Frequency distribution of misdemeanor types" = misdemeanor_types_all)
table_misdemeanortype_freq_perc <- prop.table(table_of_misdemeanortype_freq) %>% `*`(100) %>% round(2)

table_of_misdemeanortype_freq

print("Five most frequent types of misdemeanor cases by percentage")
sort(table_misdemeanortype_freq_perc,decreasing=TRUE)[1:5]
```

## Years 2012-2016

```{r, echo=FALSE}
misdemeanor_types_all2 <- unlist(strsplit(as.character(avalik2$SyndmusLiik), ","))
misdemeanor_types2 <- unique(misdemeanor_types_all2)
table_of_misdemeanortype_freq2 <- table("Frequency distribution of misdemeanor types" = misdemeanor_types_all2)
table_misdemeanortype_freq_perc2 <- prop.table(table_of_misdemeanortype_freq2) %>% `*`(100) %>% round(2)

table_of_misdemeanortype_freq2

print("Five most frequent types of misdemeanor cases by percentage")
sort(table_misdemeanortype_freq_perc2,decreasing=TRUE)[1:5]

```
# Misdemeanor cases by days
# Year 2017
```{r, echo=FALSE}
seventeenonly_group_paev <- count(seventeenonly, seventeenonly$ToimNadalapaev)

#ggplot(data=seventeenonly_group_paev, aes(x=seventeenonly_group_paev$`seventeenonly$ToimNadalapaev`, y=seventeenonly_group_paev$n)) +
#  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
#  guides(fill=FALSE) +
#  xlab("Day of week") + ylab("Number of offences") +
#  ggtitle("Comparison of misdemeanor by day")

x <- c("Esmaspäev","Teisipäev","Kolmapäev","Neljapäev","Reede","Lapäev","Pühapäev")
tr <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")

kkkk <- prop.table(table(seventeenonly$ToimNadalapaev))
names(kkkk) <- c("Monday","Wednesday","Saturday","Thursday","Sunday","Friday","Tuesday")
from_table <- data.frame(kkkk, row.names = names(kkkk))
from_table2 <- rbind(from_table[1,],from_table[7,],from_table[2,],from_table[4,],from_table[6,],from_table[3,],from_table[5,])
```

```{r, echo=FALSE}
ggplot(data= from_table2, aes(x=Var1, y=Freq )) +
   geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
   guides(fill=FALSE) +
   xlab("Day of week") + ylab("Probability of a misdemeanor happening") +
   ggtitle("Distribution of reported cases by weekday 2017") + scale_x_discrete(limits=tr)
```



# Years 2012-2016
```{r, echo=FALSE}
x <- c("Esmaspäev","Teisipäev","Kolmapäev","Neljapäev","Reede","Lapäev","Pühapäev")
tr <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")

kkkk2 <- prop.table(table(avalik2$ToimNadalapaev))
names(kkkk2) <- c("Monday","Wednesday","Saturday","Thursday","Sunday","Friday","Tuesday")
from_table3 <- data.frame(kkkk2, row.names = names(kkkk2))
from_table33 <- rbind(from_table3[1,],from_table3[7,],from_table3[2,],from_table3[4,],from_table3[6,],from_table3[3,],from_table3[5,])
```

```{r, echo=FALSE}
ggplot(data= from_table33, aes(x=Var1, y=Freq )) +
   geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
   guides(fill=FALSE) +
   xlab("Day of week") + ylab("Probability of a misdemeanor happening") +
   ggtitle("Distribution of reported cases by weekday 2012-2016") + scale_x_discrete(limits=tr)
```


# Monetary loss caused by misdemeanor
The following section displays the monetary loss caused by misdemeanor.
## 2017

```{r, echo=FALSE}
losses_2017 <- count(seventeenonly, "Loss during 2017"=Kahjusumma)
prev_losses <- count(avalik2, "Loss during 2012-2016"=avalik2$Kahjusumma)

prop.table(table("Loss"=seventeenonly$Kahjusumma))

losses_2017
```


## 2012 - 2016

```{r, echo=FALSE}
prop.table(table("Loss during 2012-2016"=avalik2$Kahjusumma))

prev_losses
```


# Places where misdemeanor took place
## 2017
```{r, echo=FALSE}
mdm_by_place_all <- unlist(strsplit(as.character(seventeenonly$KohtLiik), ","))

table_of_mdm_by_place <- table("Frequency distribution of misdemeanor cases by place" = mdm_by_place_all)

table_of_mdm_by_place_freq <- prop.table(table_of_mdm_by_place)  %>% `*`(100) %>% round(2)
table_of_mdm_by_place_freq
print("Probabilistically three most frequent places where offense might take place")
sort(table_of_mdm_by_place_freq,decreasing=TRUE)[1:3]
```

## 2012 - 2016
```{r, echo=FALSE}
mdm_by_place2_all <- unlist(strsplit(as.character(avalik2$KohtLiik), ","))

table_of_mdm_by_place2 <- table("Frequency distribution of misdemeanor cases by place" = mdm_by_place2_all)
table_of_mdm_by_place2_freq <- prop.table(table_of_mdm_by_place2)  %>% `*`(100) %>% round(2)
table_of_mdm_by_place2_freq
print("Probabilistically three most frequent places where offense might take place")
sort(table_of_mdm_by_place2_freq,decreasing=TRUE)[1:3]
```


# Misdemeanor rate by month

```{r, echo=FALSE}
a_by_ym <- count(seventeenonly, "Year and month" = substr(seventeenonly$ToimKpv,1,6))
b_by_ym <- count(avalik2, "Year and month" = substr(avalik2$ToimKpv,1,6))
all_by_ym <- rbind( a_by_ym, b_by_ym)

all_all <- rbind( seventeenonly, avalik2)
mdm_by_month <- table(substr(all_all$ToimKpv,1,7))
mdm_by_month_rate <- prop.table(mdm_by_month)
mdm_by_month_rate_table <- data.frame(mdm_by_month_rate, row.names = names(mdm_by_month_rate))
```

```{r, echo=FALSE}
ggplot(data=mdm_by_month_rate_table[1:12,], aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Year and month") + ylab("rate of offense") +
  ggtitle("Misdemeanor rate by month through 2012") +
  coord_flip()
```

```{r, echo=FALSE}
ggplot(data=mdm_by_month_rate_table[13:24,], aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Year and month") + ylab("rate of offense") +
  ggtitle("Misdemeanor rate by month through 2013") +
  coord_flip()
```

```{r, echo=FALSE}
ggplot(data=mdm_by_month_rate_table[25:36,], aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Year and month") + ylab("rate of offense") +
  ggtitle("Misdemeanor rate by month through 2014") +
  coord_flip()
```

```{r, echo=FALSE}
ggplot(data=mdm_by_month_rate_table[37:48,], aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Year and month") + ylab("rate of offense") +
  ggtitle("Misdemeanor rate by month through 2015") +
  coord_flip()
```

```{r, echo=FALSE}
ggplot(data=mdm_by_month_rate_table[49:60,], aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Year and month") + ylab("rate of offense") +
  ggtitle("Misdemeanor rate by month through 2016") +
  coord_flip()
```

```{r, echo=FALSE}
ggplot(data=mdm_by_month_rate_table[61:72,], aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Year and month") + ylab("rate of offense") +
  ggtitle("Misdemeanor rate by month through 2017") +
  coord_flip()
```

```{r, echo=FALSE}
ggplot(data=mdm_by_month_rate_table, aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Year and month") + ylab("rate of offense") +
  ggtitle("All from 2012 to 2017 for nice comparison") +
  coord_flip()
```

# Rate of misdemeanor cases by month

```{r, echo=FALSE}
mdm_by_month2 <- table(substr(avalik2$ToimKpv,6,7))
mdm_by_month_rate2 <- prop.table(mdm_by_month2)
mdm_by_month_rate_table2 <- data.frame(mdm_by_month_rate2, row.names = names(mdm_by_month_rate2))

ggplot(data=mdm_by_month_rate_table2, aes(x=Var1, y=Freq)) +
  geom_bar(colour="black", fill="#DD8888", width=.8, stat="identity") +
  guides(fill=FALSE) +
  xlab("Month") + ylab("Probability of a misdemeanor happening") +
  
  ggtitle("Distribution of reported cases by month (2012-2016)") 
```
